---
title: "Step 1: Loading & processing images"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Step 1: Loading & processing images}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  fig.align = "center",
  fig.width = 6,
  strip.white = TRUE
)
```

> How to get images into R.

Most of the actual `recolorize` functions can take a path to an image directly as input, so these steps are entirely optional. I include it here to show the option of loading the image into R first to do some basic manipulations, and because it can be useful to understand how images are structured as arrays in R.

## Reading an image into R

We can read in an image by passing the filepath to the `readImage` function. This is a pretty generic function (almost every image processing package in R has something similar); the `recolorize` version doesn't even assign the output to a special class (so don't try to print it).

```{r setup}
library(recolorize)

# define image path - we're using an image that comes with the package
img_path <- system.file("extdata/corbetti.png", package = "recolorize")

# load image
img <- readImage(img_path, resize = NULL, rotate = NULL)

# it's just an array with 4 channels:
dim(img)
```

An image is a numeric array with either 3 or 4 channels (R, G, B, and optionally alpha for transparency). JPG images will only have 3 channels; PNG images will have 4. This is quite a small image (243x116 pixels) with 4 channels.

We can plot the whole array as an image, or plot one channel at a time. Notice that the red patches are bright in the R channel, same for blue-B channel, green-G channel, etc—and that the off-white patch is bright for all channels, while the black patches are dark in all channels. The alpha channel is essentially just a mask that tells us which parts of the image to ignore when processing it further.

```{r}
layout(matrix(1:5, nrow = 1))
plotImageArray(img, main = "RGB image")
plotImageArray(img[ , , 1], main = "R channel")
plotImageArray(img[ , , 2], main = "G channel")
plotImageArray(img[ , , 3], main = "B channel")
plotImageArray(img[ , , 4], main = "Alpha channel")
```

Optionally, when you load the image, you can resize it (highly recommended for large images) and rotate it. Image processing is computationally intensive, and R is not especially good at it, so downsampling it usually a good idea. A good rule of thumb for downsampling is that you want the smallest details you care about in the image (say, spots on a ladybug) to be about 5 pixels in diameter (so if your spots have a 20 pixel diameter, you can set `resize = 0.25`).

The only other thing you might do to your images before sending them to the main `recolorize` functions is `blurImage`. This is really useful for minimizing color variation due to texture (e.g. scales on a lizard, feathers on a bird, sensory hairs on an insect), and you can apply one of several smoothing algorithms from the `imager` package, including edge-preserving blurs:

```{r}
blurred_img <- blurImage(img, blur_function = "blur_anisotropic",
                         amplitude = 10, sharpness = 0.2)
```

This step is optional: most of the `recolorize` functions will accept a path to an image as well as an image array.

## Further reading

* [Introduction](Introduction.html)  
* [Step 0: Image acquisition and preparation](step00_prep.html)  
* Step 1: Loading & processing images
* [Step 2: Initial clustering](step02_initial_cluster.html)  
* [Step 3: Refinement](step03_refinement.html)  
* [Step 4: Tweaks & edits](step04_manual_tweak.html)  
* [Step 5: Visualizing & exporting output](step05_visualization_export.html)  

> You can also tour the functions in the [function gallery](gallery.html).



